---
title: "Homework 3"
author: "Advik Ravichandran"
date: "February 13th, 2025"
output:
  pdf_document:
    toc: false
    toc_depth: '3'
  html_document:
    toc: false
    toc_float: false
    toc_depth: 3
    number_sections: false
    theme: journal
---

**Github Link:** <https://github.com/advikr/SDS315-HW-3>

# Problem 1

### Theory A

**Claim:** Gas stations charge more if they lack direct competition in sight.

**Evidence:**

```{r echo = FALSE, warning = FALSE, results = 'hide', message = FALSE}
# Loading libraries
library(ggplot2)
library(dplyr)
library(tidyverse)
library(knitr)
library(mosaic)
```

```{r echo = FALSE, warning = FALSE}

# Finding Evidence
gas_prices <- read.csv("gasprices.csv")

boot_gas_price <- do(10000)*diffmean(Price ~ Competitors, data=mosaic::resample(gas_prices))
confint(boot_gas_price, level = 0.95)
mean(Price ~ Competitors, data = gas_prices)

ggplot(gas_prices) + geom_boxplot(aes(x = Price, y = Competitors)) + 
  labs(
    y = "Competitors Present",
    title = "Gas Prices With and Without Competitors"
  )



```

When competitors are present, we are 95% confident that gas prices change between -\$0.06 and \$0.01 dollars.

**Conclusion:**

The theory is not supported by the data because the confidence interval for the difference in gas prices based on if competitors are present includes 0. We are not confident that the gas prices change due to if a competitor is present or not.

### Theory B

**Claim:** The richer the area, the higher the gas prices.

**Evidence:**

```{r echo = FALSE, warning = FALSE}
# Finding Evidence
gas_prices <- read.csv("gasprices.csv")

# Scatterplot
ggplot(gas_prices) + geom_point(aes(x = Price, y = Income)) + 
  labs(
    title = "The Correlation Between Gas Prices and Income"
  )
cor(gas_prices$Price, gas_prices$Income)

boot_gas_income <- do(10000)*coef(lm(Price ~ Income, data=mosaic::resample(gas_prices)))
confint(boot_gas_income, level = 0.95)

```

The correlation between price of gas and median household income is 0.3962, which is a moderate, positive relationship between the two variables. With 95% confidence, we can say that with every \$10,000 increase in income, there is between \$0.67 and \$1.76 increase in gas prices.

**Conclusion:**

As the evidence shows, this theory is valid and prices do increase as median income increases.

### Theory C

**Claim:** Gas stations at stoplights charge more.

**Evidence:**

```{r echo = FALSE, warning = FALSE}
ggplot(gas_prices) + geom_boxplot(aes(x = Price, y = Stoplight)) + 
  labs(
    title = "Gas Prices With or Without a Stoplight Present"
  )

boot_gas_stop <- do(10000)*diffmean(Price ~ Stoplight, data=mosaic::resample(gas_prices))
confint(boot_gas_stop, level = 0.95)
```

We are 95% confident that stoplights can change gas prices between -\$0.04 and \$0.03 dollars.

**Conclusion:**

The theory is not supported by the data because the confidence interval for the difference in gas prices based on if stoplights are present includes 0. We are not confident that the gas prices change due to if a stoplight is present or not.

### Theory D

**Claim:** Highways with direct highway access charge more.

**Evidence:**

```{r echo = FALSE, warning = FALSE}

ggplot(gas_prices) + geom_boxplot(aes(x = Price, y = Highway)) + 
  labs(
    title = "Gas Prices With and Without Highway Access"
  )

boot_gas_highway <- do(10000)*diffmean(Price ~ Highway, data=mosaic::resample(gas_prices))
confint(boot_gas_highway, level = 0.95)
```

We are 95% confident that highways can change gas prices between \$0.01 and \$0.08 dollars.

**Conclusion:**

As the evidence shows, since the confidence interval does not include 0, this theory is valid and prices do increase if there is a stoplight present.

### Theory E

**Claim:** Shell charges more than all other non-Shell brands.

**Evidence:**

```{r echo = FALSE, warning = FALSE}
gas_prices <- read.csv("gasprices.csv")
gas_shell <-
  gas_prices %>%
  mutate(isShell = Name == "Shell")
ggplot(gas_shell) + geom_boxplot(aes(x = Price, y = isShell)) + 
  labs(
    y = "Shell Owned",
    title = "Gas Prices Compared Between Shell and Non Shell Owned Stations"
  )

boot_gas_shell <- do(10000)*diffmean(Price ~ isShell, data=mosaic::resample(gas_shell))
confint(boot_gas_shell, level = 0.95)
```

When the gas station is owned by Shell, we are 95% confident that gas prices change between -\$0.01 and \$0.06 dollars.

**Conclusion:**

The theory is not supported by the data because the confidence interval for the difference in gas prices based on if Shell owns the gas station includes 0. We are not confident that the gas prices change due to if Shell owns the company or not.

# Problem 2

### Part A

```{r echo = FALSE, warning = FALSE}
cars <- read.csv("sclass.csv")

cars_2011 <- filter(cars, year == 2011, trim ==  "63 AMG")
avg_mileage <- do(10000)*mean(~mileage, data=mosaic::resample(cars_2011))

confint(avg_mileage, level = 0.95)
```

We are 95% confident that Mercedes cars 2011 year and 63 AMG trim have an average mileage between 26254 and 31773 miles.

### Part B

```{r echo = FALSE, warning = FALSE}
cars <- read.csv("sclass.csv")

cars_2014 <-
  filter(cars, year == 2014, trim == "550") %>%
  mutate(isBlack = color == "Black")

boot_2014 <- do(10000)*prop(~isBlack, data = resample(cars_2014))
confint(boot_2014, level = 0.95)
```

We are 95% confident that the proportion for Mercedes 2014 cars that are black with a 550 trim is between 41.7% and 45.3%.

# Problem 3

### Part A

```{r echo = FALSE, warning = FALSE}
shows <- read.csv("nbc_pilotsurvey.csv")

shows_a <-
  filter(shows, Show == "Living with Ed" | Show == "My Name is Earl")

means <-
  shows_a |>
  group_by(Show) |>
  summarize(mean_value = mean(Q1_Happy, na.rm = TRUE))

diffmean(Q1_Happy ~ Show, data=shows_a)

boot_shows_a = do(10000)*diffmean(Q1_Happy ~ Show, data=mosaic::resample(shows_a))
confint(boot_shows_a, level = 0.95)
```

**Question:** Between the show Living with Ed and My Name is Earl, is there evidence that one show consistenlty produces a higher mean Q1_Happy response score among viewers?

**Approach:** First, I filtered the dataset so it would only include the ratings of the two shows. Then, I calculated the difference in mean of Q1_Happy scores, and then bootstrapped it 10,000 times to reduce error caused by uncertainty. Finally, I calculated the confidence interval with the bootstrapped data.

**Results:** This bootstrap shows that we are 95% confident that My Name is Earl has a mean score of Q1 Happiness in between -0.4 and 0.1 than Living with Ed.

**Conclusion:** Our findings are insignificant because the confidence interval goes through 0, which mean that there is a chance that the means are the same. Stakeholders should know that no, there is no evidence that one show consistently produces a higher mean Q1 Happy response score among viewers.

### Part B

```{r echo = FALSE, warning = FALSE}
shows_b <- filter(shows, Show == "The Biggest Loser" | Show == "The Apprentice: Los Angeles")
diffmean(Q1_Annoyed ~ Show, data = shows_b)

means_showsb <-
  shows_b |>
  group_by(Show) |>
  summarize(mean_value = mean(Q1_Annoyed, na.rm = TRUE))

boot_b <- do(10000)*diffmean(Q1_Annoyed ~ Show, data=mosaic::resample(shows_b))
confint(boot_b, level = 0.95)
```

**Question:** Between the show The Biggest Loser and the show The Apprentice: Los Angeles, which show had the viewers the most annoyed based off of the Q1_Annoyed score?

**Approach:** First, I filtered the dataset so it would only include the ratings of the two shows. Then, I calculated the difference in mean of Q1_Annoyed scores, and then bootstrapped it 10,000 times to reduce error caused by uncertainty. Finally, I calculated the confidence interval with the bootstrapped data.

**Results:** This bootstrap shows that we are 95% confident that the show The Biggest Loser has a mean score of Q1_Annoyed 0.527 and 0.024 lower than the Apprentice: Los Angeles.

**Conclusion:** Our findings are significant because the confidence interval does not go through 0, which mean that there is no chance that the means are the same. Stakeholders should know that there is a show which had viewers the most annoyed based off of the Q1_Annoyed scores, with that show being The Biggest Loser.

### Part C

```{r echo = FALSE, warning = FALSE}
shows_c <- 
  filter(shows, Show == "Dancing with the Stars")
shows_c$IsConfusing <- ifelse(shows_c$Q2_Confusing >= 4, TRUE, FALSE)
boot_shows_c <- do(10000) * prop(~IsConfusing, data = resample(shows_c))
confint(boot_shows_c, level = 0.95)
```

**Question:** What proportion of American TV watchers would we expect to give a response of 4 or greater to the “Q2_Confusing” question?

**Approach:** First, I created a new dataset which filtered it by only showing the show of interest, and made a new column, IsConfusing, that returns TRUE or FALSE if the Q2_Confusing is greater or equal to 4. Then, I bootstrapped the data 10,000 times to reduce the error caused by uncertainty and then used that data to determine the confidence interval at a 95% level.

**Results:** This bootstrap shows that we are 95% confident that the true proportion of viewers who watched the show Dancing with the Stars found confusing falls between 3.87% and 11.60%.

**Conclusion:** Our findings are significant because the confidence interval does not go through 0. Stakeholders should know that about 4% and 11% find this show confusing, which may raise problems in the future.

# Problem 4

```{r echo = FALSE, warning = FALSE}
ebay <- read.csv("ebay.csv")

ebay_ratio <- 
  ebay |>
  mutate(revenue_ratio = rev_after / rev_before)

diffmean(revenue_ratio ~ adwords_pause, data = ebay_ratio)
boot_ebay <- do(10000)*diffmean(revenue_ratio ~ adwords_pause, data=mosaic::resample(ebay_ratio))
confint(boot_ebay, level = 0.95)
# Both CI's are negative, meaning results are significant
```

**1) Question:** Does paid search advertising on Google create extra revenue for Ebay?

**2) Approach:** I first calculated the ratio for each row in the dataset, creating a new column, and then bootstrapped it 10,000 times to reduce uncertainty levels and calculated a confidence interval of 95% level.

**3) Results:** We are 95% confident that without the paid Google search advertising, Ebay is losing between 9% and 1% of revenue.

**4) Conclusion:** Our findings are significant because the confidence interval does not go through zero(-9% and -1%), suggesting that removing the Google search advertising has had a measurable, negative impact on Ebay's revenue. Given these statistics, Ebay should carefully choose whether or not to continue using Google's search advertising because it does have a small negative impact on revenue. They will have to decide whether the costs are worth the extra revenue.
